---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rethinking);library(ggplot2);library(rstan)

data = read.csv("Assignment3Data.csv")
#NAs for alot of ADOS scores - they were only collected at first visit
#1 = ASD
#0 = TD
```

own notes: 
all questions can be answered looking only at visit 1
1) focus on asd in visit 1. How did you bulit the model, what priors are used. The quality of the model
should you scale the predictors or outcomes?



## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ (verbal, nonverbal and social) and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity (ADOS) and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS


```{r}
#make subset with only visit 1 and asd children
data2 = subset(data,Visit == "1" & ASD == "1")

model1 <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + b*VerbalIQ ,
    a ~ dnorm( 20 , 20 ) ,
    b ~ dnorm( 0 , 10 ) ,
    sigma ~ dunif( 0 , 10 )
    ) ,
  data=data2 )

precis(model1)


#plotting
plot( ADOS ~ VerbalIQ , data2 , col=col.alpha(rangi2,0.5) ) 

mu = link(model1)

# define sequence of verbalIQ to compute predictions 
# these values will be on the horizontal axis
Verbal.seq <- seq( from=5 , to=50 , by=1 )
# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( model1 , data=data.frame(VerbalIQ=Verbal.seq) )
str(mu)

# use type="n" to hide raw data
plot( ADOS ~ VerbalIQ , data2 , type="n" )
# loop over samples and plot each mu value
for ( i in 1:100 )
  points( Verbal.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.1) )

# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )#compute the mean of each coloumn of the matrix mu
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 ) 

# plot raw data
# fading out points to make line and interval more visible
plot( ADOS ~ VerbalIQ , data=data2 , col=col.alpha(rangi2,0.5) )
# plot the MAP line, aka the mean mu for each weight
lines( Verbal.seq , mu.mean )
# plot a shaded region for 89% HPDI
shade( mu.HPDI , Verbal.seq )


#Or
ggplot(data=data2, aes(VerbalIQ, ADOS)) + geom_point() + geom_smooth(method = "lm")

#__________________________
#non-verbal IQ

model2 <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + b*NonVerbalIQ ,
    a ~ dnorm( 20 , 20 ) ,
    b ~ dnorm( 0 , 10 ) ,
    sigma ~ dunif( 0 , 10 )
    ) ,
  data=data2 )

dfmodel2 = (precis(model2))
dfmodel2

#__________________________
#social iQ

model3 <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + b*SocialIQ ,
    a ~ dnorm( 20 , 20 ) ,
    b ~ dnorm( 0 , 10 ) ,
    sigma ~ dunif( 0 , 10 )
    ) ,
  data=data2 )

precis(model3)




```


2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.


```{r}

d <- read.csv("Assignment3Data.csv")
d_visit1 <- d[d$Visit == 1,]
d_visit1_ASD <- d_visit1[d_visit1$ASD == 1,]
d_visit1_ASD$ADOS_c <- d_visit1_ASD$ADOS - mean(d_visit1_ASD$ADOS) 
d_visit1_ASD$NonVerbalIQ_c <- d_visit1_ASD$NonVerbalIQ - mean(d_visit1_ASD$NonVerbalIQ)
d_visit1_ASD$VerbalIQ_c <- d_visit1_ASD$VerbalIQ - mean(d_visit1_ASD$VerbalIQ)
d_visit1_ASD$SocialIQ_c <- d_visit1_ASD$SocialIQ - mean(d_visit1_ASD$SocialIQ)

#Does it make sense to have them all in the same model? 

cor(data2$VerbalIQ, data2$NonVerbalIQ)
cor(data2$VerbalIQ, data2$SocialIQ)
cor(data2$NonVerbalIQ, data2$SocialIQ)

modelmulti <- map( 
             alist(
               ADOS_c ~ dnorm( mu , sigma ) ,
               mu <- a + bv*VerbalIQ_c + bn*NonVerbalIQ_c + bs*SocialIQ_c ,
               a ~ dnorm( 20 , 20 ) ,
               bv ~ dnorm( 0 , 10 ) ,
               bn ~ dnorm( 0 , 10 ) ,
               bs ~ dnorm( 0, 10), 
               sigma ~ dunif( 0 , 30 )
               ) ,
             data = d_visit1_ASD )
precis(modelmulti)
plot(precis(modelmulti))


#look for big inflations in sd
nonverbal_soc_model <- map( 
             alist(
               ADOS_c ~ dnorm( mu , sigma ) ,
               mu <- a  + bn*NonVerbalIQ_c + bs*SocialIQ_c ,
               a ~ dnorm( 20 , 20 ) ,
               bn ~ dnorm( 0 , 10 ) ,
               bs ~ dnorm( 0, 10), 
               sigma ~ dunif( 0 , 30 )
               ) ,
             data = d_visit1_ASD )

plot(precis(nonverbal_soc_model))
precis(nonverbal_soc_model)

#look for big inflations in sd
verbal_soc_model <- map( 
             alist(
               ADOS_c ~ dnorm( mu , sigma ) ,
               mu <- a  + bv*VerbalIQ_c + bs*SocialIQ_c ,
               a ~ dnorm( 20 , 20 ) ,
               bv ~ dnorm( 0 , 10 ) ,
               bs ~ dnorm( 0, 10), 
               sigma ~ dunif( 0 , 30 )
               ) ,
             data = d_visit1_ASD )

plot(precis(verbal_soc_model))
precis(verbal_soc_model)

```
plots for exercise 2

```{r}
#plotting




```


3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses. #own notes: does symptoms severity drive IQ? or the other way?? - fun to discuss

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


